Intentional release of native species undermines ecological stability

### 2022-10-18 rmd/manuscript.Rmd
rmd

---
title: "Intentional release of native species undermines ecological stability"
output:
  bookdown::pdf_document2:
   latex_engine: xelatex
   toc: false
  #bookdown::word_document2:
  # reference_docx: word_format.docx
csl: https://www.zotero.org/styles/pnas
knit: (function(inputFile, encoding) {
        rmarkdown::render(inputFile,
                          encoding = encoding, 
                          output_dir = "document_output")
      })
bibliography: reference.bib
fontsize: 11pt
mainfont: "Times New Roman"
header-includes:
  - \usepackage{lineno}
  - \linenumbers
---

@@

Massive releases of captive-bred native species ("intentional release") is a pervasive method to enhance wild populations of commercial and recreational species. However, such external inputs may disrupt the sensitive balance of species interactions that allows competing species to coexist, potentially compromising the long-term community stability. Here, we use theory and long-term data of stream fish communities to show that intentional release destabilizes community dynamics with a limited demographic benefit to the enhanced species. Our theory predicted that intentional release intensifies interspecific competition, facilitating competitive exclusion of unenhanced species that otherwise stablely coexist. In parallel, the excessive input of captive-bred individuals suppressed the natural recruitment of the enhanced species via intensified within-species competition. Consequently, the ecological community with intentional release is predicted to show reduced community density with unstable temporal dynamics. Consistent with this prediction, stream fish communities showed greater temporal fluctuations and fewer species richness in rivers with the intensive release of hatchery salmon - a major fishery resource worldwide. Our findings alarm that the current overreliance on intentional release may accelerate global biodiversity loss, threatening the provisioning of ecosystem services.

@@

Intentional release of captive-bred individuals is a common practice for conservation and natural resource management. However, we know little about its potential consequences for the whole ecological community. Here, we show that intentional release undermines community stability with little demographic benefit to the enhanced species. Theory and data agreed that intentional release destabilizes community dynamics by facilitating competitive exclusion while suppressing the natural recruitment of the enhanced species. The effect size of intentional release was striking in its magnitude, doubling temporal fluctuations of enhanced communities compared to those with no intentional release. Our findings point to major limitations of intentional release as a primary tool for biodiversity conservation and sustainability.

**Key words:** coexistence theory, competition, resilience, ecological modeling, fishery

@@

Human demands for natural resources are ever-increasing, such that active interventions are critical to the sustainable management of fisheries, forestry, and wildlife [@laikreCompromisingGeneticDiversity2010]. Releases of captive-bred native species ("intentional release") is a form of the efforts to enhance wild populations of diverse taxa [@snyderLimitationsCaptiveBreeding1996; @laikreCompromisingGeneticDiversity2010]. In particular, the magnitude of intentional release for commercial and recreational species is massive (reviewed in [@laikreCompromisingGeneticDiversity2010; @kitadaEconomicEcologicalGenetic2018]) -- a single program releases up to tens of millions of plant [@laikreReleaseAlienPopulations2006], invertebrate (bivalves, insects) [@ingsBumblebeesHumblePollinators2005; @haagNorthAmericanFreshwater2012], and vertebrate individuals (fish, birds) [@villanuaSanitaryRisksRedlegged2008; @kitadaEconomicEcologicalGenetic2018] into the wild annually. There is growing awareness that intentional release entails ecological risks, including the accumulation of deleterious alleles [@arakiGeneticEffectsCaptive2007], the spread of diseases in wild populations [@krkosekDecliningWildSalmon2007], and the intensification of intraspecific competition within the enhanced species [@kitadaLessonsLearnedJapanese2006]. Yet, this method is still pervasive in natural resource management owing to the significant economic benefit [@malkamakiSystematicReviewSocioeconomic2018; @davidEconomicFeasibilityIntensification2018; @rutledgeCostBenefitHatcheryreared1990; @kitadaEconomicEcologicalGenetic2018]. For example, Kitada @kitadaEconomicEcologicalGenetic2018 analyzed economic performance of release programs for 12 major fishery resources across the globe. The gross monetary yield exceeded the seed-production cost in nine species, suggesting that release programs are economically profitable in general.

Current debates, however, overlook the fact that we have rarely assessed the community-wide impact of intentional release. Species are all embedded in the complex web of interacting organisms, and the stable coexistence of competing species through density-dependent feedback underpins the emergent stability of ecological communities [@tilmanEcologicalConsequencesChanges1999; @carrollNicheFitnessDifferences2011a; @loreauDoesFunctionalRedundancy2004]. Intentional release may disrupt the sensitive balance of species interactions because it introduces unnaturally high numbers of individuals into the wild [@laikreCompromisingGeneticDiversity2010; @careyNativeInvadersChallenges2012; @kitadaEconomicEcologicalGenetic2018]. Hence, this form of species management may intervene in the ecological process that allows competing species to coexist, ultimately degrading the long-term community stability. The community-wide impact may be particularly serious in release programs for fishery resources as many of them release millions of hatchery individuals into a single water body (e.g., a watershed) every year. Evidence for this hypothesis is lacking, however.

Here, we show that intentional release undermines long-term community stability, which we define as the relative size of fluctuations in total community density over time [@tilmanEcologicalConsequencesChanges1999]. Our theory illuminates that intentional release compromises the stabilizing mechanism emerging from species niche differences. The present study further demonstrates the relevance of this general theory to natural systems by showing its congruence with Japanese stream fish communities, where \~10 million hatchery masu salmon (*Oncorhynchus masou masou*) are released annually for fisheries, recreation, and conservation purposes across the nation [@kitadaLessonsJapanMarine2020]. Once released as fry, masu salmon stay in freshwater for at least one year before sea migration and compete for resources with other stream-dwelling species. Therefore, this system serves as an excellent model to study the community-wide impact of intentional release. Our integrative approach provides strong evidence for the destabilizing effect of intentional release on ecological communities.

@@

We employed a multispecies Ricker model [@fowlerSpeciesDynamicsAlter2012] to simulate community dynamics with the intentional release of a constituent species (species 1). Specifically, the population density of species $i$ at time $t+1$, $N_{i,t+1}$, is modeled as:

@@

where $r_i$ is the intrinsic growth rate, $\alpha_{ij}$ the competition coefficient of species $j$ on species $i$, $K_i$ the carrying capacity, $R_t$ the number of released individuals for the enhancement of species 1, and $\varepsilon_{i,t}$ the species response to stochastic environmental fluctuations that follow a normal distribution $\text{Normal}(0,\sigma_{\varepsilon}^2)$. The parameter $\phi_i$ controls the relative fitness of captive-bred individuals as follows:

@@

Prior to our full-community analysis, we analyzed a two species community to understand the model behavior. The prior analysis revealed distinct parameter spaces predicting destabilizing, neutral, and stabilizing effects of intentional release on community dynamics (**Figures S1 and S2**). In general, intentional release had a neutral to destabilizing effect in a community with slow-growing species ($r_i < 1.5$). Meanwhile, for higher values of $r_i$ ($\ge 1.5$), intentional release stabilized community dynamics by dampening limit cycles or chaos. Carrying capacity and competition also played an important role in shifting the parameter spaces. According to this result, we chose `r nrow(list_param[[2]])` simulation scenarios, differing in the intrinsic growth of an enhanced species $r_1$, average competition strength $\bar{\alpha}$, carrying capacity $K$, and relative fitness $f_R$ (see **Methods**). To capture variation in species traits, intrinsic growth rates of unenhanced species and interspecific competition coefficients were drawn randomly from a uniform ($r_{i,i \ne 1} \sim \text{Unif}(0.5,2.5)$) and an exponential distribution ($\alpha_{ij,i \ne j} \sim \text{Exp}(1/\bar{\alpha})$). Under each parameter scenario, we ran `r attr(list_param[[2]], "n_sim")` time steps of `r max(list_sim[[2]]$n_rep)` independent communities (i.e., simulation replicates). Using the last `r unique(list_sim[[2]]$n_timestep)` time steps, we obtained the following summary statistics of the total community density $\sum_i^SN_i$ to examine the community-level response to intentional release: coefficient of variation (CV), number of species persist (defined as $N_i > 0.01$ for the entire simulation), temporal mean ($\mu$), and temporal SD ($\sigma$). We also calculated the temporal mean and SD for the enhanced ($N_1$) and unenhanced species ($\sum_{i,i \ne 1}^S N_i$) separately to infer underlying mechanisms.

When carrying capacity was small ($K=$ `r min(list_param[[2]]$k)`), our model predicted a destabilizing effect of intentional release on ecological communities as illustrated by increased CV with increasing numbers of releases (**Figure 1A**). This pattern stemmed mainly from the reduced mean of the total community density, and both enhanced and unenhanced species groups were responsible (**Figure 1A**). The enhanced species decreased because release induced the negative competitive effect exceeding the reproductive contribution of released individuals, confirming the previous theoretical prediction [@satakeStockingCaptivebredFish2012]. Meanwhile, interspecific competition reduced the unenhanced species, resulting in fewer persisting species (**Figure 1A**). Combined, the total community density decreased more sharply than individual species groups (**Figure 1A**). The SDs showed a similar trend, but the relationship was flatter at the community level (**Figure 1A**). Since a CV is a ratio of an SD to a mean, the steeper decline of the mean community density led to the increased CV. These patterns were qualitatively similar across ecological contexts with the low carrying capacity (robust to the changes in $r_1$, $\bar{\alpha}$, and $f_R$; **Figures S3-S6**).

The destabilizing effect emerges because intentional release affects the balance of species interactions that underpins community stability. In theory, the stable coexistence requires a niche difference that is large enough to overcome the relative difference in intrinsic competitive ability [@carrollNicheFitnessDifferences2011a]. Under this condition, competing species can grow from small populations because dominant species undergo stronger intraspecific competition [@carrollNicheFitnessDifferences2011a]. Such coexistence favors stable temporal dynamics of species-rich communities [@tilmanEcologicalConsequencesChanges1999] because it gives rise to "overyielding" [@loreauDoesFunctionalRedundancy2004; @carrollNicheFitnessDifferences2011a], i.e., total community density of a multispecies community exceeds what would be expected in a single species community ($\sum_i^S N_i > K$). However, intentional release is externally controlled, and the number of releases is not subject to density-dependent regulation. Therefore, released individuals impose additional intra- and interspecific competition that interferes with the ecological process producing overyielding.

Intentional release, however, had little or a stabilizing influence on community dynamics when carrying capacity was sufficiently large ($K=$ `r max(list_param[[2]]$k)`; **Figures S7-S10**). In particular, intentional release increased the enhanced species with a low population growth rate ($r_1=$ `r min(list_param[[2]]$r1)`; **Figure S7**), suggesting that weak intraspecific competition within the enhanced species (= $r_1 K^{-1}$) is needed for intentional release to be effective. Our result is reasonable because some of the best evidence for successful intentional release comes from re-introduction/translocation programs of extirpated species or population enhancement of long-lived endangered species [@deguchiTranslocationHandrearingResult2017; @jonesRestorationMauritiusKestrel1995; @okahisaEvaluationOngoingCrested2022; @yackulicAssessingPopulationImpacts2021]. In the Sado island (Japan), for example, the endangered Crested Ibis *Nipponia nippon* showed exponential growth of the population since the initial reintroduction of ten captive-bred individuals in 2008 [@okahisaEvaluationOngoingCrested2022]. Similarly, translocations of Humpback chub, a federally-listed fish species in the US, seem highly cost-effective in enhancing their tributary populations in the Little Colorado River [@yackulicAssessingPopulationImpacts2021]. Two common features exist in these successful cases. First, the receiving ecosystem supports suitable habitats (i.e., high carrying capacity). Second, the number of releases in these conservation-oriented programs is rather small (several tens to hundreds) [@fischerAssessmentPublishedResults2000] in comparison to release programs for commercial and recreational species (up to several tens of millions). These conditions meet the theoretical assumptions with which intentional release led to increased abundance of the enhanced species. However, the preponderance of unsuccessful reintroduction programs implies that these conditions are not always sufficed [@fischerAssessmentPublishedResults2000; @haagNorthAmericanFreshwater2012; @strayerEssayMakingMost2019]. Indeed, non-negligible numbers of projects seem to release captive-bred individuals into the wild without knowing whether the release habitat is suitable for the focal species [@fischerAssessmentPublishedResults2000; @haagNorthAmericanFreshwater2012]. Our analysis suggests that pre-release examination and restoration of environmental capacity may be a key to successful release programs with minimal impacts on other community members.

@@

To demonstrate the relevance of our general theory to natural systems, we assessed the potential impacts of the intentional release of masu salmon (**Figure 1E**) on the long-term stability of stream fish communities in Hokkaido, Japan. In the protected watersheds (all separated by the ocean; **Figure 1C, D**), a long-term program exists to monitor stream fish communities along with the official release records. The release of masu salmon began in 1950s, with the annual average ranging from `r round(min(df_stock_mu$mean_stock), 2)` (no release) to `r round(max(df_stock_mu$mean_stock), 2)` million fish across watersheds during 1999 -- 2019. The majority of release occurs in spring at fish survey locations, after which salmon fry stay in freshwater for at least one year with other stream fishes (see **Figure S16** for a co-occurrence matrix in our dataset). Therefore, the study system sets the stage for a "natural experiment" to test our theoretical predictions. We used the data from 1999 to 2019 at `r n_distinct(df_fish$site_id)` sites within `r n_distinct(df_fish$river)` independent watersheds (see **Methods** for selection criteria). Using hierarchical Bayesian models, we quantified the effect of intentional release on community dynamics while accounting for potential effects of climates and local abiotic factors.

As predicted, stream fish communities showed greater temporal fluctuations (higher CV) at sites with intensive release (**Figure 1B**). The effect was striking in its magnitude, almost doubling the CV at the highest release level. Our analysis strongly supported the positive relationship between the CV and the number of "effective" releases (**Figure 2**; see **Methods** for the Bayesian estimation of effective release), in which the probability of the regression coefficient being positive was `r op(df_beta %>% filter(response == "cv") %>% pull(prob), 2)` (**Table S6**). This pattern was associated with the reduced long-term average of the total community density and fewer taxonomic richness (**Figures 1B and 2**), and both enhanced (masu salmon) and unenhanced fish groups contributed to this trend (**Figures 1B and 2;** see **Tables S6-S8** for full statistics). In the meantime, the SDs had vague relationships with intentional release (**Figures 1B and 2**).

To gain better insights into the driving mechanisms, we fitted a multi-species Ricker model to a subset of our time-series data (nine sites with few missing observations and no intentional release; **Methods**). The parameter estimates of this model are directly comparable to those used in the theoretical Ricker model. The estimated competition coefficients $\alpha_{ij}$ ranged `r op(min(df_pd$alpha_prime), 2)` -- `r op(max(df_pd$alpha_prime), 2)` with the median of `r op(median(df_pd$alpha_prime), 2)` (**Figure 3**), suggesting that competition is a likely mechanism as our theory assumes (see **Figure S17** for all pairwise estimates). Our statistical inference is supported by previous studies. For example, interference competition between masu salmon and other salmonid species is well documented [@miyasakaFoodHabitDivergence2003; @hasegawaComparisonCompetitiveAbility2004]. Diets of salmonid species are innately overlapped (e.g., drift terrestrial insects), such that they interfere with each other for profitable foraging spots [@hasegawaComparisonCompetitiveAbility2004]. Exploitative competition between masu salmon and benthic fish (e.g., sculpins and stone loaches in our study system) is also likely because subordinate masu salmon in the dominance hierarchy consumes benthic macroinvertebrates [@satoStagespecificFunctionalResponses2014], which are dominant prey items for many benthic fish species [@natsumedaEvaluationEcologicalFeatures2010; @natsumedaDietFoodSelectivity2012]. Importantly, hatchery masu salmon may intensify intra- and interspecific competition in the wild [@hasegawaWildMasuSalmon2018; @sahashiEvaluationContributionHatchery2015; @hasegawaPulsedSuppliesSmall2022]. Hatchery salmons are larger and more aggressive than wild individuals of a comparable life stage, increasing the likelihood of intense intra- and interspecific competition [@weberInteractionsHatcheryWild2003]. Indeed, field and experimental studies confirmed that hatchery masu salmon competed with wild masu salmon and other stream fishes during their freshwater life stage [@hasegawaWildMasuSalmon2018; @sahashiEvaluationContributionHatchery2015; @hasegawaPulsedSuppliesSmall2022].

The reduced fitness of hatchery masu salmon may also play a role in the observed response. Like other salmonids, masu salmon exhibits a life history of partial migration [@moritaLatitudinalVariationGrowth2010; @sahashiMigrationCostsDrive2013]. In our study region, the majority of individuals migrate to the ocean and spend approximately one year before spawning, although some males stay in freshwater for their entire life [@moritaLatitudinalVariationGrowth2010; @moritaTemperaturedependentVariationAlternative2014]. It has been shown that ocean-migrating adults of hatchery masu salmon show lower return rates to the spawning river [@sahashiWildGenesBoostinpress], and such genetic effects may exacerbate exponentially as the number of generations in captivity increases [@arakiGeneticEffectsCaptive2007]. Typically, hatchery masu salmon is reared in captivity for two to four generations; thus, the fitness decline may be substantial. In addition, the aggressive behavior of hatchery salmons may make them vulnerable to predation [@weberInteractionsHatcheryWild2003]. The multiple fitness disadvantages may add to negative density dependence to influence community stability.

Importantly, socioeconomic factors control the number of releases. The hatchery program aims to produce and release a constant number of hatchery masu salmon every year, although uncontrollable factors (e.g., diseases, budget allocation) generate considerable variation in the release number among years. Therefore, the number of releases is determined without accounting for the current conditions of the recipient ecosystem. As such, released fish are probably "excessive" and may cause resource competition that would otherwise not exist.

>>>>

---
title: "Intentional release of native species undermines ecological stability"
output:
  bookdown::pdf_document2:
   latex_engine: xelatex
   toc: false
   number_sections: false
  #bookdown::word_document2:
  # reference_docx: word_format.docx
csl: https://www.zotero.org/styles/pnas
knit: (function(inputFile, encoding) {
        rmarkdown::render(inputFile,
                          encoding = encoding, 
                          output_dir = "document_output")
      })
bibliography: reference.bib
fontsize: 11pt
mainfont: "Times New Roman"
header-includes:
  - \usepackage{lineno}
  - \linenumbers
---

@@

Massive releases of captive-bred native species ("intentional release") is a pervasive method to enhance wild populations of commercial and recreational species. However, such external inputs may disrupt the sensitive species interactions that allow competing species to coexist, potentially compromising long-term community stability. Here, we use theory and long-term data of stream fish communities to show that intentional release destabilizes community dynamics with a limited demographic benefit to the enhanced species. Our theory predicted that intentional release intensifies interspecific competition, facilitating the competitive exclusion of unenhanced species that otherwise stably coexist. In parallel, the excessive input of captive-bred individuals suppressed the natural recruitment of the enhanced species via intensified within-species competition. Consequently, the ecological community with the intentional release is predicted to show reduced community density with unstable temporal dynamics. Consistent with this prediction, stream fish communities showed greater temporal fluctuations and fewer species richness in rivers with the intensive release of hatchery salmon - a major fishery resource worldwide. Our findings alarm that the current overreliance on intentional release may accelerate global biodiversity loss, threatening the provisioning of ecosystem services.

@@

The intentional release of captive-bred individuals is a common practice for conservation and natural resource management. However, we know little about its potential consequences for the whole ecological community. Here, we show that the intentional release undermines community stability with little demographic benefit to the enhanced species. Theory and data agree that intentional release destabilizes community dynamics by facilitating competitive exclusion while suppressing the natural recruitment of the enhanced species. The effect size of the intentional release was striking in its magnitude, doubling temporal fluctuations of enhanced communities compared to those with no intentional release. Our findings point to major limitations of intentional release as a primary tool for biodiversity conservation and sustainability.

**Keywords:** coexistence theory, competition, resilience, ecological modeling, fishery

@@

Human demands for natural resources are ever-increasing, such that active interventions are critical to the sustainable management of fisheries, forestry, and wildlife [@laikreCompromisingGeneticDiversity2010]. Releases of captive-bred native species ("intentional release") is a form of the efforts to enhance wild populations of diverse taxa [@snyderLimitationsCaptiveBreeding1996; @laikreCompromisingGeneticDiversity2010]. In particular, the magnitude of intentional release for commercial and recreational species is massive (reviewed in [@laikreCompromisingGeneticDiversity2010; @kitadaEconomicEcologicalGenetic2018]) -- a single program releases up to tens of millions of plant [@laikreReleaseAlienPopulations2006], invertebrate (bivalves, insects) [@ingsBumblebeesHumblePollinators2005; @haagNorthAmericanFreshwater2012], and vertebrate individuals (fish, birds) [@villanuaSanitaryRisksRedlegged2008; @kitadaEconomicEcologicalGenetic2018] into the wild annually. There is growing awareness that intentional release entails ecological risks, including the accumulation of deleterious alleles [@arakiGeneticEffectsCaptive2007], the spread of diseases in wild populations [@krkosekDecliningWildSalmon2007], and the intensification of intraspecific competition within the enhanced species [@kitadaLessonsLearnedJapanese2006]. Yet, this method is still pervasive in natural resource management owing to the significant economic benefit [@malkamakiSystematicReviewSocioeconomic2018; @davidEconomicFeasibilityIntensification2018; @rutledgeCostBenefitHatcheryreared1990; @kitadaEconomicEcologicalGenetic2018]. For example, Kitada @kitadaEconomicEcologicalGenetic2018 analyzed the economic performance of release programs for 12 major fishery resources across the globe. The gross monetary yield exceeded the seed-production cost in nine species, suggesting that release programs are economically profitable in general.

Current debates, however, overlook the fact that we have rarely assessed the community-wide impact of intentional release. Species are all embedded in the complex web of interacting organisms, and the stable coexistence of competing species through density-dependent feedback underpins the emergent stability of ecological communities [@tilmanEcologicalConsequencesChanges1999; @carrollNicheFitnessDifferences2011a; @loreauDoesFunctionalRedundancy2004]. Intentional release may disrupt the sensitive balance of species interactions because it introduces unnaturally high numbers of individuals into the wild [@laikreCompromisingGeneticDiversity2010; @careyNativeInvadersChallenges2012; @kitadaEconomicEcologicalGenetic2018]. Hence, this form of species management may intervene in the ecological process that allows competing species to coexist, ultimately degrading the long-term community stability. Evidence for this hypothesis is lacking, however.

Here, we show that intentional release undermines long-term community stability, which we define as the relative size of fluctuations in total community density over time [@tilmanEcologicalConsequencesChanges1999]. Our theory illuminates that intentional release compromises the stabilizing mechanism emerging from species niche differences. The present study further demonstrates the relevance of this general theory to natural systems by showing its congruence with Japanese stream fish communities, where \~10 million hatchery masu salmon (*Oncorhynchus masou masou*) are released annually for fisheries, recreation, and conservation purposes across the nation [@kitadaLessonsJapanMarine2020]. Once released as fry, masu salmon stay in freshwater for at least one year before sea migration and compete for resources with other stream-dwelling species. Therefore, this system serves as an excellent model for studying the community-wide impact of intentional release. Our integrative approach provides strong evidence for the destabilizing effect of intentional release on ecological communities.

@@

We employed a multi-species Ricker model [@fowlerSpeciesDynamicsAlter2012] to simulate community dynamics with the intentional release of a constituent species (species 1). Specifically, the population density of species $i$ at time $t+1$, $N_{i,t+1}$, is modeled as:

@@

where $r_i$ is the intrinsic growth rate, $\alpha_{ij}$ the competition coefficient of species $j$ on species $i$ ($\alpha_{ii} = 1$), $K_i$ the carrying capacity, $R_t$ the number of released individuals for the enhancement of species 1, and $\varepsilon_{i,t}$ the species response to stochastic environmental fluctuations that follow a normal distribution $\text{Normal}(0,\sigma_{\varepsilon}^2)$. The parameter $\phi_i$ controls the relative fitness of captive-bred individuals as follows:

@@

Prior to our full-community analysis, we analyzed a two-species community to understand the model behavior. The prior analysis revealed distinct parameter spaces predicting destabilizing, neutral, and stabilizing effects of intentional release on community dynamics (**Figures S1 and S2**). In general, the intentional release had a neutral to destabilizing effect in a community with slow-growing species ($r_i < 1.5$). Meanwhile, for higher values of $r_i$ ($\ge 1.5$), intentional release stabilized community dynamics by dampening limit cycles or chaos. Carrying capacity and competition also played an important role in shifting the parameter spaces. According to this result, we chose `r nrow(list_param[[2]])` simulation scenarios, differing in the intrinsic growth of an enhanced species $r_1$, average competition strength $\bar{\alpha}$, carrying capacity $K$, and relative fitness $f_R$ (see **Methods**). To capture variation in species traits, intrinsic growth rates of unenhanced species and interspecific competition coefficients were drawn randomly from a uniform ($r_{i,i \ne 1} \sim \text{Unif}(0.5,2.5)$) and an exponential distribution ($\alpha_{ij,i \ne j} \sim \text{Exp}(1/\bar{\alpha})$). Under each parameter scenario, we ran `r attr(list_param[[2]], "n_sim")` time steps of `r max(list_sim[[2]]$n_rep)` independent communities (i.e., simulation replicates). Using the last `r unique(list_sim[[2]]$n_timestep)` time steps, we obtained the following summary statistics of the total community density $\sum_i^SN_i$ to examine the community-level response to intentional release: coefficient of variation (CV), number of species persist (defined as $N_i > 0.01$ for the entire simulation), temporal mean ($\mu$), and temporal SD ($\sigma$). We also calculated the temporal mean and SD for the enhanced ($N_1$) and unenhanced species ($\sum_{i,i \ne 1}^S N_i$) separately to infer underlying mechanisms.

When carrying capacity was small ($K=$ `r min(list_param[[2]]$k)`), our model predicted a destabilizing effect of intentional release on ecological communities as illustrated by increased CV with increasing numbers of releases (**Figure 1A**). This pattern stemmed mainly from the reduced mean of the total community density, and both enhanced and unenhanced species groups were responsible (**Figure 1A**). The enhanced species decreased because release induced the negative competitive effect exceeding the reproductive contribution of released individuals, as reported in a previous theoretical study [@satakeStockingCaptivebredFish2012]. Meanwhile, interspecific competition reduced the unenhanced species, resulting in fewer persisting species (**Figure 1A**). Combined, the total community density decreased more sharply than individual species groups (**Figure 1A**). The SDs showed a similar trend, but the relationship was flatter at the community level (**Figure 1A**). Since a CV is a ratio of an SD to a mean, the steeper decline of the mean community density led to the increased CV. These patterns were qualitatively similar across ecological contexts with the low carrying capacity (robust to the changes in $r_1$, $\bar{\alpha}$, and $f_R$; **Figures S3-S6**).

The destabilizing effect emerges because intentional release affects the balance of species interactions that underpins community stability. In theory, stable coexistence requires a niche difference that is large enough to overcome the relative difference in intrinsic competitive ability [@carrollNicheFitnessDifferences2011a]. Under this condition, competing species can grow from small populations because dominant species undergo stronger intraspecific competition [@carrollNicheFitnessDifferences2011a]. Such coexistence favors stable temporal dynamics of species-rich communities [@tilmanEcologicalConsequencesChanges1999] because it gives rise to "overyielding" [@loreauDoesFunctionalRedundancy2004; @carrollNicheFitnessDifferences2011a], i.e., total community density of a multi-species community exceeds what would be expected in a single species community ($\sum_i^S N_i > K$). However, the intentional release is externally controlled, and the number of releases is not subject to density-dependent regulation. Therefore, released individuals impose additional intra- and interspecific competition that interferes with the ecological process producing overyielding.

The intentional release had little or a stabilizing influence on community dynamics when carrying capacity was sufficiently large ($K=$ `r max(list_param[[2]]$k)`; **Figures S7-S10**). In particular, intentional release increased the enhanced species with a low population growth rate ($r_1=$ `r min(list_param[[2]]$r1)`; **Figure S7**), suggesting that weak intraspecific competition within the enhanced species (= $r_1 K^{-1}$) is needed for intentional release to be effective. The weak-competition scenario is likely in re-introduction programs of extirpated species or population enhancement of long-lived endangered species. In fact, some of the best evidence for successful intentional release comes from re-introduction efforts [@deguchiTranslocationHandrearingResult2017; @jonesRestorationMauritiusKestrel1995; @okahisaEvaluationOngoingCrested2022; @yackulicAssessingPopulationImpacts2021]. In the Sado island (Japan), for example, the once-extirpated Crested Ibis *Nipponia nippon* showed exponential growth of the population since the initial re-introduction of ten captive-bred individuals in 2008 [@okahisaEvaluationOngoingCrested2022]. Similarly, translocations of Humpback chub, a federally-listed fish species in the US, seem highly cost-effective in enhancing their tributary populations in the Little Colorado River [@yackulicAssessingPopulationImpacts2021]. However, the preponderance of unsuccessful re-introduction programs implies that the theoretical requirements for successful intentional release are not always met [@fischerAssessmentPublishedResults2000; @haagNorthAmericanFreshwater2012; @strayerEssayMakingMost2019]. Indeed, non-negligible numbers of projects seem to release captive-bred individuals into unsuitable habitats with compromised carrying capacities [@fischerAssessmentPublishedResults2000; @haagNorthAmericanFreshwater2012]. Our analysis suggests that such programs could rather impair biodiversity without noticeable demographic benefit to the enhanced species. Pre-release examination and restoration of environmental capacity may be key to successful release programs with minimal impacts on other community members.

@@

To demonstrate the relevance of our general theory to natural systems, we assessed the potential impacts of the intentional release of masu salmon (**Figure 1E**) on the long-term stability of stream fish communities in Hokkaido, Japan. In the protected watersheds (all separated by the ocean; **Figure 1C, D**), a long-term program exists to monitor stream fish communities along with the official release records. The release of masu salmon began in the 1950s, with the annual average ranging from `r round(min(df_stock_mu$mean_stock), 2)` (no release) to `r round(max(df_stock_mu$mean_stock), 2)` million fish across watersheds during 1999 -- 2019. The majority of release occurs in spring at fish survey locations, after which salmon fry stays in freshwater for at least one year with other stream fishes (see **Figure S16** for a co-occurrence matrix in our dataset). Therefore, the study system sets the stage for a "natural experiment" to test our theoretical predictions. We used the data from 1999 to 2019 at `r n_distinct(df_fish$site_id)` sites within `r n_distinct(df_fish$river)` independent watersheds (see **Methods** for selection criteria). Using hierarchical Bayesian models, we quantified the effect of intentional release on community dynamics while accounting for the potential effects of climates and local abiotic factors.

As predicted, stream fish communities showed greater temporal fluctuations (higher CV) at sites with the intensive release (**Figure 1B**). The effect was striking in its magnitude, almost doubling the CV at the highest release level. Our analysis strongly supported the positive relationship between the CV and the number of "effective" releases (**Figure 2**; see **Methods** for the Bayesian estimation of effective release), in which the probability of the regression coefficient being positive was `r op(df_beta %>% filter(response == "cv") %>% pull(prob), 2)` (**Table S6**). This pattern was associated with the reduced long-term average of the total community density and fewer taxonomic richness (**Figures 1B and 2**), and both enhanced (masu salmon) and unenhanced fish groups contributed to this trend (**Figures 1B and 2;** see **Tables S6-S8** for full statistics). In the meantime, the SDs had vague relationships with the intentional release (**Figures 1B and 2**).

To gain better insights into the driving mechanisms, we fitted a multi-species Ricker model to a subset of our time-series data (nine sites with few missing observations and no intentional release; **Methods**). The parameter estimates of this model are directly comparable to those used in the theoretical Ricker model. The estimates of interspecific competition coefficients $\alpha_{ij}$ ranged `r op(min(df_pd$alpha_prime), 2)` -- `r op(max(df_pd$alpha_prime), 2)` with the median of `r op(median(df_pd$alpha_prime), 2)` (**Figure 3**), suggesting that competition is a likely mechanism as our theory assumes (see **Figure S17** for all pairwise estimates). Our statistical inference is supported by previous studies. For example, interference competition between masu salmon and other salmonid species is well documented [@miyasakaFoodHabitDivergence2003; @hasegawaComparisonCompetitiveAbility2004]. Diets of salmonid species are innately overlapped (e.g., drift terrestrial insects), such that they interfere with each other for profitable foraging spots [@hasegawaComparisonCompetitiveAbility2004]. Exploitative competition between masu salmon and benthic fish (e.g., sculpins and stone loaches in our study system) is also likely because subordinate masu salmon in the dominance hierarchy consumes benthic macroinvertebrates [@satoStagespecificFunctionalResponses2014] that many benthic fish species prey on [@natsumedaEvaluationEcologicalFeatures2010; @natsumedaDietFoodSelectivity2012]. Importantly, hatchery masu salmon may intensify intra- and interspecific competition in the wild [@hasegawaWildMasuSalmon2018; @sahashiEvaluationContributionHatchery2015; @hasegawaPulsedSuppliesSmall2022]. Hatchery salmons are larger and more aggressive than wild individuals of a comparable life stage, increasing the likelihood of intense intra- and interspecific competition [@weberInteractionsHatcheryWild2003]. Indeed, field and experimental studies confirmed that hatchery masu salmon competed with wild masu salmon and other stream fishes during their freshwater life stage [@hasegawaWildMasuSalmon2018; @sahashiEvaluationContributionHatchery2015; @hasegawaPulsedSuppliesSmall2022].

The reduced fitness of hatchery masu salmon may also play a role in the observed response. Like other salmonids, masu salmon exhibits a life history of "partial migration"; in our study region, the majority of individuals migrate to the ocean and spend approximately one year before spawning, although some males stay in freshwater for their entire life [@moritaLatitudinalVariationGrowth2010; @sahashiMigrationCostsDrive2013; @moritaTemperaturedependentVariationAlternative2014]. It has been shown that ocean-migrating adults of hatchery masu salmon show lower return rates to the spawning river [@sahashiWildGenesBoostinpress], and such genetic effects may accumulate as the number of generations in captivity increases [@arakiGeneticEffectsCaptive2007]. Typically, hatchery masu salmon are reared in captivity for two to four generations. Thus, the fitness decline may be substantial. In addition, the aggressive behavior of hatchery salmons may make them vulnerable to predation [@weberInteractionsHatcheryWild2003]. The multiple fitness disadvantages may add to negative density dependence to influence community stability.

External control of intentional release is a theoretical premise for the destabilizing effect to emerge, and this is exactly how the release of masu salmon is operated. The hatchery program aims to produce and release a constant number of hatchery masu salmon every year, although uncontrollable factors (e.g., disease, budget allocation) generate some variation in the release number between years. Therefore, the number of releases is determined without accounting for the current condition of the recipient ecosystem. As such, released fish are probably "excessive" and may cause resource competition that would otherwise not exist.

It is noteworthy that the magnitude of the release program at the protected watersheds (annual average maximum = `r op(max(df_stock_mu$mean_stock), 2)` million fish) is comparable or even smaller when compared with those for other commercial and recreational species [@laikreCompromisingGeneticDiversity2010; @kitadaEconomicEcologicalGenetic2018]. For example, 0.5 -- 1.2 million juveniles of Black sea bream are released annually in Kagoshima bay, Japan, where substantial declines in genetic diversity and wild individuals have been reported [@kitadaRigorousMonitoringLargescale2019]. Similarly, three to four millions of Red-legged partridges are introduced repeatedly for sports hunting in rural Spain [@villanuaSanitaryRisksRedlegged2008]. Therefore, our results should not be viewed as an extreme scenario. We also must mention that a meta-analysis focusing on fishery resources revealed that only a few massive release programs led to an increased abundance of the enhanced species [@arakiHatcheryStockingHelp2010]. Yet, none of those studies statistically quantified the community-level impact of massive releases. Future studies addressing this knowledge gap are therefore highly valuable.
